---
title: "Module 2, Assignment 2"
author: "Ellen Bledsoe"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Details

### Purpose

The goal of this assignment is to assess your ability to write custom functions and iterate code using a `for loop`. 

### Task

Write R code which produces the correct answers and correctly interpret the plots produced.

### Criteria for Success

-   Code is within the provided code chunks
-   Code is commented with brief descriptions of what the code does
-   Code chunks run without errors
-   Code produces the correct result
    -   Code that produces the correct answer will receive full credit
    -   Code attempts with logical direction will receive partial credit
-   Written answers address the questions in sufficient detail

### Due Date

October 11 at midnight MST

# Assignment Questions

For this assignment, we will continue using the `penguins` data set from the `palmerpenguin` data package. 

You should reference the [2_Functions]() and [3_SickFish]() lessons to complete this assignment.

1. Load both the `tidyverse` and `palmerpenguin` packages into our workspace.
```{r}
library(tidyverse)
library(palmerpenguins)
```

### Optional
If you want the `penguins` data frame to show up in the environment, run the following code chunk.
```{r}
penguins <- penguins
```
Right now, the bill length, bill depth, and flipper length measurements are in millimeters (we know this from the column names). A colleague has asked for the data in centimeters. 

2. First, let's write a custom function that will convert any value we give it from millimeters to centimeters. Add in all the necessary information, including:
* argument(s)
* code to convert from mm to cm (hint: divide by 10)
* what value the function should return.
```{r}
mm_to_cm <- function(mm){
  cm <- mm / 10
  return(cm)
}
```

Don't forget to run the code chunk above to add your function to the environment!

