---
title: "Module 1, Assignment 1: Getting to Know the Team"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- no need to change; uses the current date when rendered
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description

### Purpose

The goal of this assignment is to get comfortable using R to look at 1- and 2-dimensional data sets.

### Task

Write R code to successfully answer each question below.

### Criteria for Success

-   Code is within the provided code chunks
-   Code chunks run without errors
-   Code will be assessed as follows:
    -   Produces the correct answer using the requested approach: 100%
    -   Generally uses the right approach, but a minor mistake results in an incorrect answer: 90%
    -   Attempts to solve the problem and makes some progress using the core concept, but returns the wrong answer and does not demonstrate comfort with the core concept: 50%
    -   Answer demonstrates a lack of understanding of the core concept: 0%
-   Written answers address the questions in sufficient detail

### Due Date

Jan 25 at 11am MST

# Assignment Questions

Each question is worth 2 points unless otherwise specified.

## Vectors

Let's start working with vectors, or 1-dimensional data, first.

Run this chunk of code to create a vector of penguins counts. Once you run the code, the object `counts` should appear in the environment.

```{r vector}
# vector with counts of penguins
counts <- c(2, 9, 4, 3, 6, 7, 1, 0, 3)
```

1.  What data class is `counts`? Use a function to find out.

```{r data_class}

```

2.  Write a line of code that pulls out the number 9 from the `counts` vector.

```{r second_value}

```

3.  Calculate the average number of penguins that were counted.

```{r avg_penguins}

```

## Data Frames

Now that we've practiced with vectors, let's move on to 2-dimensional data.

Remember those fun/silly questions about our trip to Antarctica? It's time to play around with that data!

The following code chunk will read the data into RStudio. Be sure to run it before try to answer the questions! A bunch of stuff will pop up, but it should work just fine. Once you run the code, you will see a dataframe called `team_data` show up in the environment.

```{r import_data}
team_data <- read.csv("team_antarctica.csv")
```

4.  Using either a function or looking in the environment, answer the following questions (2 points each).

-   What are the dimensions of the data (how many rows and how many columns)?
-   What are the names of the columns in the data?
-   What is the data *class* for each column?

*Answer:*

```{r}
# using code is optional
```

5.  Take a look at the data frame using the `head()` function. Typically, `head()` provides the first 6 rows of data. Modify one of the arguments in `head()` so that the line of code prints the first 10 rows. (If you aren't sure how to do that, remember how you can look for help about functions!)

```{r examine_data}

```

6.  Using what you know about sub-setting data frames, write a line of code the pulls out the parka color for Unique ID 4. (Hint: count the columns!)

```{r parka7}

```

When we have a large data set like this, it is often helpful to summarize the data in some way. The next few questions will help use get a better understanding of the content of the data set.

7.  On average, how did people rate their ability to be in a remote location?

```{r avg_remote_location}

```

8.  What are the minimum and maximum distances that would be traveled by a team member to get to Antarctica? Use the `min()` and `max()` functions.

```{r min_max_distance}

```

## Bonus (up to 2 points)!

What animal should be on our team flag?

First, create a vector called `mascot` that has only the values from the `team_flag` column. Next, use the `table()` function on the `mascot`. This will give you the number of times each option was chosen. According to the results, which animal should be on our team flag?

*Answer:*

```{r flag_mascot}

```

# Turning in Your Assignment

Follow these steps to successfully turn in your assignment on D2L.

1.  Click the `Knit` button up near the top of this document. This should produce a PDF file that shows up in the `Files` panel on the bottom-right of your screen.
2.  Click the empty box to the left of the PDF file.
3.  Click on the blue gear near the top of the `Files` panel and choose Export.
4.  Put your last name at the front of the file name when prompted, then click the Download button. The PDF file of your assignment is now in your "Downloads" folder on your device.
5.  Head over to D2L and navigate to Module 1 Assignment 1. Submit the PDF file that you just downloaded.
