---
title: "Module 1 Assignment 3: Getting to Know your Home"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description

### Purpose

The goal of this assignment is to get comfortable using the `tidyverse` with 2-dimensional data sets and compare this process to using base R.

### Task

Write R code using the `tidyverse` to successfully answer each question below.

### Criteria for Success

-   Code is within the provided code chunks
-   Code is commented with brief descriptions of what the code does
-   Code chunks run without errors
-   Code produces the correct result
-   This is the one time I *will* take points off for not using `tidyverse`...

### Due Date

Sept 15 at midnight MDT

# Assignment Questions

For this final assignment for Module 1, you'll be working with another real-world data set--a collection of data from climate stations scattered across Antarctica.

1.  Load in the `tidyverse` package.

```{r load_tidyverse}
# load packages
library(tidyverse)
```

2.  Load in the data file (called aggregated_station_data.csv). Save the data as an object called `weather`.

(Hint: remember to use the `tidyverse` function for reading in data! If it doesn't pop up when you start typing, you probably forgot to run the code chunk above).

```{r load_data}
# read in weather station data
weather <- read_csv("../data/aggregated_station_data.csv")
```

3.  How many rows and columns are in the data? Type your answers below:

    rows: 139, 144

    columns: 12

4.  Write code to determine how many unique stations are in the data set.

(Hint: look up the help file for the `distinct()` function).

```{r unique_stations}
# number of unique stations
weather %>% 
  distinct(station_id) %>% 
  count() # optional 

```

5.  Write a chunk of code that calculates the mean temperature for the month of January for each station.

    Some hints:

    -   take note of how months are represented in the data

    -   think about using the pipe, how we choose which rows we want, and how we split-apply-combine

```{r mean_jan_temp}
weather %>% 
  filter(month == 1) %>% 
  group_by(station_id) %>% 
  summarize(mean_temp = mean(temp))
```

6.  Write code to determine which station had the coldest average temperature in January.

(hint: look up the `arrange()` function)

```{r}
weather %>% 
  filter(month == 1) %>% 
  group_by(station_id) %>% 
  summarize(mean_temp = mean(temp)) %>% 
  arrange(mean_temp)
```
