---
title: "Module 1, Assignment 2: Getting to Know the Team"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: 
  html_document: default    # <- most common output style
  github_document: default
  pdf_document: default 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description

### Purpose
The goal of this assignment is to get more comfortable dealing with large 
2-dimensional data sets that you'll often see in the wild. 

### Task
Write R code to successfully answer each question below.

### Criteria for Success
* Data and packages are loaded 
* Code is within the provided code chunks
* Code is commented with brief descriptions of what the code does
* Code chunks run without errors 
* Code produces the correct result
* Questions requiring writing are answered fully

### Due Date


# Assignment Questions

We need to add some new members to Team Antarctica! We have a dataset with 1000
people who have qualified to join the mission and some general data about them.
In this assignment, you'll be exploring this dataset to learn more about your
incoming colleagues.

The data we are using are in a file called "mission_team_data.csv" and should
already be in your data folder.

_Remember to comment your code and run each chunk of code as you go!_

1. In your own words, describe what a "package" is (1-2 sentences).


2. Load the tidyverse package into the workspace.

```{r load_tidyverse, message = FALSE}
# load the tidyverse
library(tidyverse)
```

3. Import the dataset. Assign the dataset to an object called team_data. Use the
read_csv() function to do this.
```{r import_data}
team_data <- read_csv("../data/mission_team_data.csv")
```

4. Running the code chunk above produces a message that gives us some useful
information, even before we look at the dataset. 
* What are the dimensions of the data?
* What are the names of the columns in the data?
* What data _type_ do you expect to find in each column (i.e., numbers, character strings, T/F, etc.)

5. Now, let's look at the data. Uses the glimpse() _and_ head() functions to examine the data. 
```{r examine_data}
glimpse(team_data)
head(team_data)
```

When we have a large dataset like this, it is often helpful to summarize the data
in some way. The next few questions will help use get a better understanding of 
the content of the dataset.

6. Write some code to determine how many of our new colleagues are male, female, and intersex (biological sex
is not binary, and not really even a spectrum but more of a mosaic! Happy to chat about this more if you're interested, but lets just focus on getting some numbers for now).
```{r count_sexes}

```

7. What is the mean height (cm) and mean weight (lbs) of the incoming team members?
Add comments in your code chunk with the answers.

```{r avg_height_weight}

```

8. What is the most common specialty for incoming team members? Least common?
```{r specialties}

```
