---
title: "Module 1, Assignment 2: Getting to Know the Team"
author: "Ellen Bledsoe"     # <- your name here!
date: "`r Sys.Date()`"      # <- uses the current date when rendered
output: 
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description

### Purpose
The goal of this assignment is to get comfortable using base R to look at 1- and 2-dimensional datasets. 

### Task
Write R code to successfully answer each question below.

### Criteria for Success
* Code is within the provided code chunks
* Code is commented with brief descriptions of what the code does
* Code chunks run without errors 
* Code produces the correct result

### Due Date


# Assignment Questions

_Remember to comment your code and run each chunk of code as you go!_

## Vectors 

Let's start working with vectors, or 1-dimensional data, first.

Run this chunk of code to create a vector of data to use.
```{r vector}
# vector with counts of penguins
counts <- c(2, 9, 4, 3, 6, 7, 1, 0, 3)
```

1. What data class is `counts`? Write a line of code that tells you.
```{r data_class}
# data class for counts
class(counts)
```

2. Write a line of code that pulls out the 2nd value in the vector.
```{r second_value}
# second value of counts vector
counts[2]
```

3. Calculate the average number of penguins that were counted.
```{r avg_penguins}
# mean number of penguins
mean(counts)
```

## Data Frames

Now that we've practiced with vectors, let's move on to 2-dimensional data.

We need to add some new members to Team Antarctica! We have a dataset with 1000
people who have qualified to join the mission and some general info about them.
In this part of the assignment, you'll be exploring this dataset to learn more 
about your incoming colleagues.

The data we are using are in a file called "mission_team_data.csv" and should
already be in your data folder for Module 1. The following code chunk will read 
in the data. Be sure to run it before try to answer the questions!

```{r import_data}
team_data <- read.csv("../data/mission_team_data.csv")
```

Running the code chunk above produces a message that gives us some useful
information, even before we look at the dataset. 
* What are the dimensions of the data?
* What are the names of the columns in the data?
* What data _class_ do you expect to find in each column (i.e., numbers, character strings, T/F, etc.)

4. Take a look at the data frame using the `head()` function. Typically, `head()`
provides the first 6 rows of data. Modify one of the arguments in `head()` so 
that the line of code prints the first 10 rows. (If you aren't sure how to do
that, remember how you can look for help about functions!)
```{r examine_data}
head(team_data, 10)
```

5. Using what you know about subsetting data frames, write a line of code the 
pulls out the age for your new colleague, Marissa Yu (row 3).
```{r}
team_data[3, 4]
```

When we have a large dataset like this, it is often helpful to summarize the data
in some way. The next few questions will help use get a better understanding of 
the content of the dataset.

6. What is the average weight (in kg) of the incoming team members? 
```{r average_weight}
mean(team_data$weight_kg)
```


7. What is the mean height (cm) and mean weight (lbs) of the incoming team members?
Add comments in your code chunk with the answers.

```{r min_max_height}
# minimum height
min(team_data$height_cm)

# maximum height
max(team_data$height_cm)
```

8. Create a data frame that only includes rows of data for people whose specialty
is "Management." 
```{r }
team_data[team_data$specialties == "Management",]
```


9. How many different specialties are represented amongst your new colleauges?
This will require two steps.
    a. First, use the `unique()` function to pull out each distinct value from the
  specialties column. Save this as an object called `specialties`, which is now
  a vector.
    b. Use the `length()` function to count how many distinct specialties are in
  the `specialties` vector.

```{r specialties}
specialties <- unique(team_data$specialties)
specialties
length(specialties)
```
