---
title: "Module 3-1"
author: "Keaton Wilson"
date: "2/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to the Problem and Summarizing Data  

## The Challenge  

Given the issues with our farmed fish, we will need to supplement our diets with
some fish that are wild-caught for a while. These fish are caught by members of 
Team Antarctica, but there is quite a bit of risk involved.

One of the main hazards to these fishing teams are leopard seals. They're apex 
predators in this ecosystem and can cause serious injury, especially when large 
schools of fish are involved (a researcher was killed by a Leopard Seal in 
Antarctica in 2003).  

One way we're working to remove some of this danger is by working with data from
large mammal researchers. They have radio collared on a number of seals that live
in areas that could be fished. Fishing boats are equipped with radios that can
detect the presence of seals in the area and avoid high-risk sites.  

In recent months, a problem has arisen. Some of the collars are failing, leading
to some very close calls! While attempting to untangle a net, one team member was
pulled into the water when a seal lunged at some fish trapped in the net.  
 
The large mammal team replaces collars frequently based on the two manufacturers' 
recommendations for battery life and general wear and tear, but it seems as if 
some of the collars are dying earlier than expected and putting our team in danger. 

We've been tasked with determining why the units are failing and if we can tie 
it to a particular manufacturer.

## The Data  

Our data on collars is in a file called "collar_data.csv."  

### Explore the Data
1. Two manufacturers  
2. Battery life - The average number of days a particular collar lasts (this is
recorded in the unit and stored when the battery dies)  
3. Signal distance - the maximum signal distance that a particular collar was 
recorded at.  
4. Fail - collars that have failed in the past (e.g. they've been recovered 
from seals that were noticed by the team but that didn't ping the radio 
equipment).  

### Group Challenge 1 - Summarizing

Let's practice 

* summarizes the data by calculating the mean, min and max for:
  * battery life 
  * signal distance, while 
* counts up the number of failures
* by manufacturer

```{r summarize}
library(tidyverse)

collars <- read_csv("../data/collar_data.csv")
glimpse(collars)

bat_life_and_signal <- collars %>%
  mutate(fail = as.logical(fail), 
         maker = as.factor(maker)) %>%
  dplyr::select(-collar_id) %>%
  group_by(maker) %>%
  summarize(bat_life_mean = mean(battery_life),
            bat_life_min = min(battery_life),
            bat_life_max = max(battery_life),
            signal_mean = mean(signal_distance),
            signal_min = min(signal_distance),
            signal_max = max(signal_distance)) 

fail_count <- collars %>%
  group_by(maker) %>%
  summarize(fail_count = sum(fail))

final <- bind_cols(bat_life_and_signal, 
                   fail_count %>% dplyr::select(-maker))

```

<!---maybe this is where we introduce pivot_wider and pivot_longer and data table structure?--->
<!---maybe better introduced in Module 2, although pivoting is important for ggplot2,
so it isn't completely out of place here in Module 3--->


## Visualizing the Data - an introduction to a better way to visualize the data
So far, we've used the base R plotting syntax, but there is another way. It's 
part of the tidyverse and its from a package called ggplot2.  

Instead of explaining the syntax to you, we're going to do an in-class reading 
assignment.  

[Article to read](https://towardsdatascience.com/a-comprehensive-guide-to-the-grammar-of-graphics-for-effective-visualization-of-multi-dimensional-1f92b4ed4149)  

Spend 10 minutes reading the article above (take notes!). When the time is up, 
I'll signal you to get into your groups and discuss the article. There is a lot 
of material here - make sure that you cover what the grammar of graphics is, 
what the different components are, and how they relate to code structure and 
syntax. We'll be iteratively building some ggplot code on the data above, and 
I'll be asking groups to explain each part as we go - so be prepared!  

```{r}
ggplot(collars, aes(y = signal_distance, x = battery_life, col = maker)) +
  geom_point() +
  xlab("Battery Life") +
  ylab("Signal Distance") +
  scale_color_discrete(name = "Collar Maker")
```

