---
title: 'Module 1: Intro to Coding in R'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Coding

## Learning Outcomes

-   Students will be able to define the following terms:
    -   object
    -   assignment
    -   vector
    -   function
    -   data frame
-   Students will be able to run code line-by-line and as code chunks
    from an Rmarkdown file.
-   Students will be able to comment their code effectively.
-   Students will be able to write code assign values to variables and
    use these variables to perform various operations.
-   Students will be able to use help files to learn how to use
    functions.
-   Students will be able to recall and explain how functions operate,
    and the basic syntax around functions (arguments, auto-completion,
    parentheses).
-   Students will be able to differentiate different data classes in R.
-   Students will learn how to create their own data structures
    (vectors, data frames).

## Assigning Objects

Assignments are really key to almost everything we do in R. This is how
we create permanence in R. Anything can be saved to an object, and we do
this with the assignment operator, `<-`.

The short-cut for `<-` is `Alt + -` (or `Option + -` on a Mac)

```{r}
# Assigning Objects
mass <- 47.5            # this is the mass in kg
age <- 122
mass <- mass * 2        # multiply
age <- age - 20         # subtract
mass_index <- mass/age  # divide
mass_sq <- mass^2       # raise to an exponent

# This is simple and you'll rarely do it in real-world scenarios. 

```

## 1-Dimensional Data: Vectors

We can also assign more complex group of elements of the same type to a
particular object. This is a basic data structure in R.

```{r}
mass_kg <- c(3, 2, 4, 9, 7, 3, 6)
mass_kg

animals <- c("cat", "rat", "bat")
animals
# R does everything in vectors
```

## Data classes

There are a few main types in R, and they behave differently.

-   numeric (numbers)
    -   integer (no decimals allowed)
    -   double (decimals allowed--interchangeable with numeric)
-   character (letters or mixture)
-   logical (True or False; T or F)
-   factors (best used for ordinal data; levels indicate the order)

```{r}
# Examples of different data classes
mass_kg     # numeric, integer, double
animals     # character

animal_size <- as.factor(c("small", "medium", "large"))
animal_size  # factor, put in order [ordinal data]

logic <- c(T, F, F, T)  # logical
logic
```

Vectors have to contain elements that are all of the same class.

```{r, eval = FALSE}
vec <- (1, 1.000, "1", TRUE)
```

### Sub-setting Vectors

Sometimes we want to pull out and work with specific values from a
vector. This is called sub-setting (taking a smaller set of the
original)

```{r}
# Use square brackets
mass_kg[2]
mass_kg[2:4]
```

## Functions

Functions are pre-written bits of codes that perform specific tasks for
us.

```{r}
## Functions

# functions are always followed by parentheses
# anything you type into the parentheses are called arguments
mass_kg_mean <- mean(mass_kg)   # average of the mass_kg vector from above
mass_kg_mean

round(mass_kg_mean)             # rounding
round(mass_kg_mean, digits = 2) # round to 2 digits past 0

# to get more information about a function, use the help function
help(round) # or type ?help
```

### Group Challenge

Let's practice! Write a few lines of code that do the following:

-   create a vector with a number from 6 to 1
-   assign the vector to an object named `vec`
-   subset `vec` to include the last 3 numbers (should include 3, 2, 1)
-   find the sum of the numbers (hint: use the `sum()` function)

```{r}
vec <- c(6, 5, 4, 3, 2, 1)
vec

vec <- vec[4:6]
vec

sum(vec)
```

Already finished? See if you can condense your code down any further or
turn around and help out a neighbor.

```{r}
vec <- seq(6, 1)
sum(vec[4:6])
```

## 2-Dimensional Data: Data Frames

Most of the data you will encounter is two-dimensional, i.e., it has
columns and rows. Its structure resembles a spreadsheet. R is really
good with these types of data.

-   **rows** go side-to-side
-   **columns** go up-and-down

Data frames are made up of multiple vectors. Each vector becomes a
column.

```{r}
# Create a simple data frame
plants <- data.frame(height = c(55, 17, 42, 47, 68, 39),
                     nitrogen = c("Y", "N", "N", "Y", "Y", "N"))

plants
```

### Sub-setting Data Frames

Because data frames are two-dimensional, we can subset data in different
ways. We can select specific columns, specific rows, or filter rows by
values.

R always takes information for the row first, then the column.

```{r}
# Sub-setting data frames
# 2-dimensional, so you need to specify row and then column
# plants[3] # doesn't work
plants[4,1]
plants[,2]

# We can also choose specific columns using `$`
plants$height

```

### Discussion Point

This is a simple data set, but let's come up with some questions.

Example: height of plants treated with nitrogen vs. those not treated.

```{r}
# Example: height of plants treated with nitrogen vs. those not treated. 

# filter rows based on values in the nitrogen column
plants[plants$nitrogen == "Y", ]
mean(plants[plants$nitrogen == "Y", 1])
```

### Group Challenge (5 min)

Using help files on functions

As a group, find the standard deviation `sd()` of the height of plants
treated with nitrogen and those not treated with nitrogen. Which group
has the larger standard deviation?

```{r}
sd(plants[plants$nitrogen == "Y", 1])
sd(plants[plants$nitrogen == "N", 1])
```

Come up with a definition of standard deviation (Google is your
friend!), use the help file to find out how the `sd()` function works,
and be prepared to show the code you used.

## Helpful Functions

Below are some functions that I often find very helpful when working
with vectors and data frames:

-   `str()`
-   `head()` and `tail()`
-   `length()`
-   `ncol()` and `nrow()`
-   `names()`

```{r}
str(plants)

head(plants)
head(plants, n = 4)
tail(plants, n = 4)

length(plants)
length(plants$height)

ncol(plants)
nrow(plants)

names(plants)
```

# Rmarkdown and Code Chunks (10 minutes)

Rmarkdown (.Rmd) is a file format that let's us incorporate text and
code into one document seamlessly. In fact, it is the file format for
this document!

To open a new Rmarkdown file, go the File \> New File \> R Markdown.
Enter a title for the document (short but descriptive), your name, and
make sure the default output HTML is selected.

When the Rmd file opens, you'll see that there is already stuff in it!
The file will have a header with the information you entered, some code
chunks, and some text explaining how Rmd files work.

-   For writing text, you can type as you would normally
-   Code chunks are a bit different. To include text in them, you will
    need to put a \# in front. R will not read anything after the \# as
    code. Chunks look like this:

```{r}

```

Type your R code in between the lines with the \`\`\`. A quick shortcut
for adding a code chunk is Ctrl + Atl + i (Cmd + Opt + i on a Mac).
Alternatively, you can go to Code \> Insert Chunk.

To run a chunk of code, click the green arrow on the far right side of
the chunk.

Let's work with an example code chunk:

```{r}
# My first code chunk!
# Ellen Bledsoe
# Aug 2022

# Examining a pre-built data set in R
Orange

# printing maximum circumference of the biggest tree
max(Orange$circumference)

```

### Group challenge

Construct a new Rmarkdown script (with all of the appropriate formatting
we discussed) that calculates the average (mean) circumference of trees
in the orange data set. Save the mean as an object called "avg_circ."

Save the script and be ready to share it.

If you're already finished, try this challenge. Find the average age for
a subset of trees (3-5).
