---
title: 'Module 1: Intro to Coding in R'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Coding

## 2-Dimensional Data: Data Frames

Most of the data you will encounter is two-dimensional, i.e., it has
columns and rows. Its structure resembles a spreadsheet. R is really
good with these types of data.

-   **rows** go side-to-side
-   **columns** go up-and-down

Data frames are made up of multiple vectors. Each vector becomes a
column.

```{r}
# Create a simple data frame
plants <- data.frame(height = c(55, 17, 42, 47, 68, 39),
                     nitrogen = c("Y", "N", "N", "Y", "Y", "N"))

plants
```

### Sub-setting Data Frames

Because data frames are two-dimensional, we can subset data in different
ways. We can select specific columns, specific rows, or filter rows by
values.

R always takes information for the row first, then the column.

```{r}
# Sub-setting data frames
# 2-dimensional, so you need to specify row and then column
# plants[3] # doesn't work
plants[4,1]
plants[,2]

# We can also choose specific columns using `$`
plants$height

```

### Discussion Point

This is a simple data set, but let's come up with some questions.

Example: height of plants treated with nitrogen vs. those not treated.

```{r}
# Example: height of plants treated with nitrogen vs. those not treated. 

# filter rows based on values in the nitrogen column
plants[plants$nitrogen == "Y", ]
mean(plants[plants$nitrogen == "Y", 1])
```

### Group Challenge (5 min)

Using help files on functions

As a group, find the standard deviation `sd()` of the height of plants
treated with nitrogen and those not treated with nitrogen. Which group
has the larger standard deviation?

```{r}
sd(plants[plants$nitrogen == "Y", 1])
sd(plants[plants$nitrogen == "N", 1])
```

Come up with a definition of standard deviation (Google is your
friend!), use the help file to find out how the `sd()` function works,
and be prepared to show the code you used.

## Helpful Functions

Below are some functions that I often find very helpful when working
with vectors and data frames:

-   `str()`
-   `head()` and `tail()`
-   `length()`
-   `ncol()` and `nrow()`
-   `names()`

```{r}
str(plants) # structure of the object

head(plants) # first 6 values or rows
head(plants, n = 4) # first n values or rows
tail(plants, n = 4) # last n values or rows

length(plants)  # for a dataframe, number of columns
length(plants$height) # for a column, number of rows

ncol(plants)  # number of columns
nrow(plants)  # number of rows

names(plants) # list of column or object names
```

# Rmarkdown and Code Chunks (10 minutes)

Rmarkdown (.Rmd) is a file format that let's us incorporate text and
code into one document seamlessly. In fact, it is the file format for
this document!

To open a new Rmarkdown file, go the File \> New File \> R Markdown.
Enter a title for the document (short but descriptive), your name, and
make sure the default output HTML is selected.

When the Rmd file opens, you'll see that there is already stuff in it!
The file will have a header with the information you entered, some code
chunks, and some text explaining how Rmd files work.

-   For writing text, you can type as you would normally
-   Code chunks are a bit different. To include text in them, you will
    need to put a \# in front. R will not read anything after the \# as
    code. Chunks look like this:

```{r}

```

Type your R code in between the lines with the \`\`\`. A quick shortcut
for adding a code chunk is Ctrl + Atl + i (Cmd + Opt + i on a Mac).
Alternatively, you can go to Code \> Insert Chunk.

To run a chunk of code, click the green arrow on the far right side of
the chunk.

Let's work with an example code chunk:

```{r}
# My first code chunk!
# Ellen Bledsoe
# Aug 2022

# Examining a pre-built data set in R
Orange

# printing maximum circumference of the biggest tree
max(Orange$circumference)

```

### Group challenge

Construct a new Rmarkdown script (with all of the appropriate formatting
we discussed) that calculates the average (mean) circumference of trees
in the orange data set. Save the mean as an object called "avg_circ."

Save the script and be ready to share it.

If you're already finished, try this challenge. Find the average circumference of 
trees aged 484 days.
```{r}
trees484 <- Orange[Orange$age == 484, 3]
mean(trees484)
```

