---
title: 'Module 1: Intro to Coding in R'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Coding

## Learning Outcomes

-   Students will be able to define the following terms:
    -   object
    -   assignment
    -   vector
    -   function
    -   data frame
-   Students will be able to run code line-by-line and as code chunks
    from an Rmarkdown file.
-   Students will be able to comment their code effectively.
-   Students will be able to write code assign values to variables and
    use these variables to perform various operations.
-   Students will be able to use help files to learn how to use
    functions.
-   Students will be able to recall and explain how functions operate,
    and the basic syntax around functions (arguments, auto-completion,
    parentheses).
-   Students will be able to differentiate different data classes in R.
-   Students will learn how to create their own data structures
    (vectors, data frames).

## Assigning Objects

Assignments are really key to almost everything we do in R. This is how
we create permanence in R. Anything can be saved to an object, and we do
this with the assignment operator, `<-`.

The short-cut for `<-` is `Alt + -` (or `Option + -` on a Mac)

```{r}
# Assigning Objects
mass <- 47.5            # this is the mass in kg
age <- 122
mass <- mass * 2        # multiply
age <- age - 20         # subtract
mass_index <- mass/age  # divide
mass_sq <- mass^2       # raise to an exponent

# This is simple and you'll rarely do it in real-world scenarios. 

```

## 1-Dimensional Data: Vectors

We can also assign more complex group of elements of the same type to a
particular object. This is a basic data structure in R.

```{r}
mass_kg <- c(3, 2, 4, 9, 7, 3, 6)
mass_kg

animals <- c("cat", "rat", "bat")
animals
# R does everything in vectors
```

## Data classes

There are a few main types in R, and they behave differently.

-   numeric (numbers)
    -   integer (no decimals allowed)
    -   double (decimals allowed--interchangeable with numeric)
-   character (letters or mixture)
-   logical (True or False; T or F)
-   factors (best used for ordinal data; levels indicate the order)

```{r}
# Examples of different data classes
mass_kg     # numeric, integer, double
animals     # character

animal_size <- as.factor(c("small", "medium", "large"))
animal_size  # factor, put in order [ordinal data]

logic <- c(T, F, F, T)  # logical
logic
```

Vectors have to contain elements that are all of the same class.

```{r, eval = FALSE}
vec <- (1, 1.000, "1", TRUE)
```

### Sub-setting Vectors

Sometimes we want to pull out and work with specific values from a
vector. This is called sub-setting (taking a smaller set of the
original)

```{r}
# Use square brackets
mass_kg[2]
mass_kg[2:4]
```

## Functions

Functions are pre-written bits of codes that perform specific tasks for
us.

Functions are always followed by parentheses. Anything you type into the
parentheses are called arguments

```{r}
## Functions
mass_kg_mean <- mean(mass_kg)   # average of the mass_kg vector from above
mass_kg_mean

round(mass_kg_mean)             # rounding
round(mass_kg_mean, digits = 2) # round to 2 digits past 0
```

To get more information about a function, use the `help()` function.

```{r}
help(round) # or type ?help
```

We can use a function called `class()` to figure out the data type of a
vector.

```{r}

```

### Group Challenge

Let's practice! Write a few lines of code that do the following:

-   create a vector with a number from 6 to 1
-   assign the vector to an object named `vec`
-   subset `vec` to include the last 3 numbers (should include 3, 2, 1)
-   find the sum of the numbers (hint: use the `sum()` function)

```{r}
vec <- c(6, 5, 4, 3, 2, 1)
vec

vec <- vec[4:6]
vec

sum(vec)
```

Already finished? See if you can condense your code down any further or
turn around and help out a neighbor.

```{r}
vec <- seq(6, 1)
sum(vec[4:6])
```

