---
title: "Module 1, Introduction to the `tidyverse`"
author: "Keaton Wilson, Ellen Bledsoe"
date: "12/4/2019, revised August 2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2-dimensional Data and the `tidyverse`

## Learning Outcomes

-   Students will be able to load packages, use help files, and navigate file paths.
-   Students will be able to use some functions of the tidyverse: select, filter, mutate, summarize and the pipe.
-   Students will compare and contrast base R and tidyverse methodology for sub-setting data frames.
-   Students will able to use tidyverse functions to summarize real-world data.

## The `tidyverse`: What is it?

Different programming languages have different syntax (language structure). The `tidyverse` is a package (more accurately, a set of packages) offered in R that all have similar goals and a unified syntax designed to work particularly well with 2-dimensional data.

Until now, all of the coding we have done is in the original R language, which is often called "base R." The syntax in the `tidyverse` is often pretty different from base R. Both are useful, and many people often combine them, which is why we start with base R.

That said, we will be primarily using the `tidyverse` for the rest of the semester. With the exception of Module 1 Assignment 3, which is exclusively about `tidyverse`, I will never punish you for using base R in place of `tidyverse`, as long as you get the same answer!

### Wait, what is a package??

Packages are one of the neatest things of working in an open-source environment like R! They contains bits of code (often in the form of functions) that can be reused, making them a core component of reproducible data science. Anyone can develop a package, and there are thousands of them doing all sorts of things.

### Explore the `tidyverse`

Group Challenge and Discussion (10 min)

Let's begin by scoping out what is in the `tidyverse`. Go to [www.tidyverse.org](www.tidyverse.org) and browse the site. In your groups, answer the 3 questions below, and be prepared to report out at the end:

1.  What are the core packages in the tidyverse?
2.  Give a one sentence summary about what each package does.
3.  What is one question you have about the tidyverse/core packages?

## Practice with the `tidyverse`

### Download and install

In most scenarios, you will need to download a package from the internet onto your computer before you can use it in RStudio. However, with RStudio Cloud, I've already done this step for you!

For future reference, though:

-   you usually only need to go through this process once until you update R
-   we use the function `install.packages()` to download the package

```{r install_tidyverse, message = FALSE}
# download and install the tidyverse package(s)
# to run the line of code, remove the # in front of the line below and run this chunk
# install.packages("tidyverse")
```

### Load into R

Any time we open R/RStudio and want to use functions from the `tidyverse`, we need to "load" the package. We use the `library()` function to do this.

When you run this code, you'll see a message that says "Attaching packages" and "Conflicts." Don't panic!

-   The first bit tells us that the core packages have been brought into our R session.
-   The "conflict" part is a little more complicated but we don't need to worry too much about it.
    -   If you're curious, though, it is telling us that there are some functions in the `tidyverse` that have the same names as functions that are automatically installed with R and that the `tidyverse` versions of those functions will be the ones that get used by default unless we specify otherwise.

```{r load_tidyverse}
# load the tidyverse (tell RStudio we want to use this package in this session)
library(tidyverse)
```

### Climate Data

To learn about the `tidyverse` syntax, we're going to use a real data set on climate change from Berkeley, CA, USA. It outlines temperatures in major cities across the world since 1750.

```{r climate_data}
# read in the data file

# `read_csv()` is part of the `tidyverse` and gives us nice options when reading in data
climate_df <- read_csv("../data/GlobalLandTemperaturesByMajorCity.csv")

# Let's take a look at the climate data. 
climate_df
```

This looks a bit different from the 2D data were were looking at before.

The `tidyverse` converts 2D data into something called a tibble! For our intents and purposes, it is basically the same as a data frame (and I'll probably call it a data frame, in reality).

This data set clearly has a lot of data! The `tidyverse` offers a great function called `glimpse()` that lets us take a quick look at the dataset.

```{r glimpse_data, options(climate_df = 1)}
# explore the data set
glimpse(climate_df)
```

## `select()`ing columns

Let's use our first function, `select()`.

```{r select}
# select allows you pick out specific columns from your data

# remember how do this with base R?
climate_df$dt
climate_df[,1:2]

# select does this but with more power 
# first argument is the data frame, then the columns
select(climate_df, dt)

# multiple columns
select(climate_df, dt, City, Country)
select(climate_df, dt:Country, -City)

# `slice` is the function that does something similar for rows
# similar process for rows, though we typically don't have row names in R
slice(climate_df, 1:4)
```

## `filter()`ing rows

`filter()` allows you filter rows by certain conditions. Recall that we did this a bit with base R.

```{r, evaluate=FALSE}

# base R
climate_df[climate_df$Country == "United States",]

# filter
filter(climate_df, Country == "United States")

# easy to write multiple conditions and to chain stuff together
filter(climate_df, Country == "United States" & AverageTemperature > 25)

# worth noting here that we haven't saved any of this. We need to write to a new object. 
us_df <- filter(climate_df, Country == "United States")

```

## Let's practice using `select()` and `filter()`

Work with the climate data we've been using this class period. Construct a small set of code that does the following:

1. Slims down the full dataframe to one that contains the columns dt, AverageTemperature and City.
2. Filters the data for Paris with an average temperature less than 22.
3. Name this new dataframe "cold_paris"

```{r}
# not piped
slim <- select(climate_df, dt, AverageTemperature, City)
filtered <- filter(slim, City == "Paris", 
                  AverageTemperature < 22)
cold_paris <- filtered
```

## The pipe

You can use the pipe operator to chain tidyverse functions together. 

This has a twofold benefit - it removes the clutter of creating a lot of intermediate objects in your workspace, but also makes things very human-readable.

```{r}
climate_df %>%
  select(dt, AverageTemperature, City) %>%
  filter(City == "Paris", AverageTemperature < 22)
```

## Learning new verbs - Group Assignment

You've had an introduction to the tidyverse verbs, but there a few more to learn. We're going to practice extracting information from the web, which is a key tool in the toolkit of data scientists.

Figure out what the `mutate()` and `summarize()` functions do. Use the web (particularly useful links include the tidyverse website, and a help-forum for stats and programming called stack overflow). Here are the specifics: 1) Be able to briefly describe each function\
2) Come up with an example from the climate data frame we've been working on.\
As always, be ready to present to your peers.
