---
title: 'Module 1: Intro to Coding in R'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
# Introduction to Coding  

## Learning Outcomes  
* Students will be able to define the following terms:
  - object
  - assignment
  - vector
  - function
  - data frame
* Students will be able to run code line-by-line and as code chunks from an
Rmarkdown file. 
* Students will be able to comment their code effectively.  
* Students will be able to write code assign values to variables 
and use these variables to perform various operations.  
* Students will be able to use help files to learn how to use functions.  
* Students will be able to recall and explain how functions operate, and the 
basic syntax around functions (arguments, auto-completion, parentheses).  
* Students will be able to differentiate different data classes in R.  
* Students will learn how to create their own data structures (vectors, data frames).  


## Assigning Objects 

Assignments are really key to almost everything we do in R. This is how we create
permanence in R. Anything can be saved to an object, and we do this with the 
assignment operator, `<-`. 

The short-cut for `<-` is `Alt + -` (or `Option + -` on a Mac)

```{r}
# Assigning Objects
mass <- 47.5            # this is the mass in kg
age <- 122
mass <- mass * 2        # multiply
age <- age - 20         # subtract
mass_index <- mass/age  # divide
mass_sq <- mass^2       # raise to an exponent

# This is simple and you'll rarely do it in real-world scenarios. 

```

## 1-Dimensional Data: Vectors
We can also assign more complex group of elements of the same type to a particular
object. This is a basic data structure in R.
```{r}
mass_kg <- c(3, 2, 4, 9, 7, 3, 6)
mass_kg

animals <- c("cat", "rat", "bat")
animals
# R does everything in vectors
```

## Data classes
There are a few main types in R, and they behave differently. 

- numeric (numbers)
  * integer (no decimals allowed) 
  * double (decimals allowed--interchangeable with numeric)
- character (letters or mixture) 
- logical (True or False; T or F) 
- factors (best used for ordinal data; levels indicate the order)   

```{r}
# Examples of different data classes
mass_kg     # numeric, integer, double
animals     # character

animal_size <- as.factor(c("small", "medium", "large"))
animal_size  # factor, put in order [ordinal data]

logic <- c(T, F, F, T)  # logical
logic
```

Vectors have to contain elements that are all of the same class.
```{r, eval = FALSE}
vec <- (1, 1.000, "1", TRUE)
```


### Subsetting Vectors
Sometimes we want to pull out and work with specific values from a vector. This is
called sub-setting (taking a smaller set of the original)
```{r}
# Use square brackets
mass_kg[2]
mass_kg[2:4]
```

## Functions

Functions are pre-written bits of codes that perform specific tasks for us.

```{r}
## Functions

# functions are always followed by parentheses
# anything you type into the parentheses are called arguments
mass_kg_mean <- mean(mass_kg)   # average of the mass_kg vector from above
mass_kg_mean

round(mass_kg_mean)             # rounding
round(mass_kg_mean, digits = 2) # round to 2 digits past 0

# to get more information about a function, use the help function
help(round) # or type ?help
```


## 2-Dimensional Data: Data Frames
Most of the data you will encounter is two-dimensional, i.e., it has columns and rows.
Its structure resembles a spreadsheet. R is really good with these types of data.  

Data frames are made up of multiple vectors. Each vector becomes a column.

```{r}
# Create a simple data frame
plants <- data.frame(height = c(55, 17, 42, 47, 68, 39),
                     nitrogen = c("Y", "N", "N", "Y", "Y", "N"))

plants
```
 
### Subsetting Data Frames
Because data frames are two-dimensional, we can subset data in different ways. We
can select specific columns, specific rows, or filter rows by values.

```{r}
# Sub-setting data frames
# 2-dimensional, so you need to specify row and then column
plants[4,1]
plants[,2]

# We can also choose specific columns using `$`
plants$height

```


### Discussion point
This is a simple data set, but let's come up with some questions. 

Example: height of plants treated with nitrogen vs. those not treated. 
```{r}
# filter rows based on values in the nitrogen column
plants[plants$nitrogen == "Y", ]
mean(plants[plants$nitrogen == "Y", 1])
```

### Group Challenge (5 min)
Using help files on functions

As a group, find the standard deviation `sd()` of the height of plants treated with 
nitrogen and those not treated with nitrogen. Which group has the larger standard
deviation? 

Come up with a definition of standard deviation, use the help file to find out how
the function works, and be prepared to show the code you used.  


# Rmarkdown and Code Chunks (10 minutes)

Rmarkdown (.Rmd) is a file format that let's us incorporate text and code into one
document seamlessly. In fact, it is the file format for this document! 

To open a new Rmarkdown file, go the File > New File > R Markdown. Enter a title
for the document (short but descriptive), your name, and make sure the default
output HTML is selected. 

When the Rmd file opens, you'll see that there is already stuff in it! The file
will have a header with the information you entered, some code chunks, and some 
text explaining how Rmd files work. 
* For writing text, you can type as you would normally
* Code chunks are a bit different. To include text in them, you will need to put
a # in front. R will not read anything after the # as code. Chunks look like this:
```{r}

```

Type your R code in between the lines with the ```. A quick shortcut for adding
a code chunk is Ctrl + Atl + i (Cmd + Opt + i on a Mac). Alternatively, you can
go to Code > Insert Chunk.

To run a chunk of code, click the green arrow on the far right side of the chunk.

Let's work with an example code chunk:
```{r}
# My first code chunk!
# Ellen Bledsoe
# Aug 2022

# Examining a pre-built data set in R
Orange

# printing maximum circumference of the biggest tree
max(Orange$circumference)

```

### Group challenge 

Construct a new Rmarkdown script (with all of the appropriate formatting we discussed) 
that calculates the average (mean) circumference of trees in the orange data set. 
Save the mean as an object called "avg_circ."

Save the script and be ready to share it.