---
title: "Module 3 Assignment 1"
author: "Ellen Bledsoe" # <- your name here!
date: '2022-10-27'      # <- the date here!
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Details

### Purpose

The goal of this assignment is to assess your ability to compare means numerically, visually, and statistically

### Task

Write R code which produces the correct answers and correctly interpret the results of visualizations and statistical tests.

### Criteria for Success

-   Code is within the provided code chunks
-   Code chunks run without errors
-   Code produces the correct result
    -   Code that produces the correct answer will receive full credit
    -   Code attempts with logical direction will receive partial credit
-   Written answers address the questions in sufficient detail

### Due Date

Oct 23 at 9:30 am MST

# Assignment Questions

In this assignment, we're going to explore another data set on wind turbines that generate a significant portion of the energy for us down here in Antarctica.

### Set-Up

Let's load the `tidyverse` and read in the data set. Call the data `turbines`.

```{r}

```

1.  Explore the data set, either through the environment or through code. Answer the following questions (2 point):

    a.  How many turbine companies are represented in the data?
    b.  What do the data in each row of data represent?

    ```{r}
    # optional; only if you want space for coding
    ```

### Numeric

2.  Generate a summary of the data set that calculates the average required [wind speed]{.underline} and mean [power output]{.underline} for each wind turbine company. Save this as a data frame called `turbine_summary`. (2 point)

```{r}

```

### Visual

3.  Create a multiple histogram plot for the [power output]{.underline} variable. (3 points)

    -   be sure to have a histogram for each turbine producer; the color and/or the fill should be determined by the maker of the turbine. They should also be transparent and not stacked vertically.
    -   set the number of bins to 10
    -   add in vertical lines for the mean values in the same color as the turbine makers (remember to reference the correct data frame!)
    -   make sure the x-axis, y-axis, and legend labels are capitalized and easier to understand
    -   use the `theme_classic()` function

```{r}

```

4.  Generate a box-and-whisker plot that compares the required [wind speed]{.underline} between different turbine makers (3 points).

    The plot should:

    -   have capitalized and more descriptive axis labels (hint: wind speed is measured in kilometers per hour---km/hr)
    -   show raw data points in addition to the boxes. The points should be jittered.
    -   use the `theme_classic()` function

```{r}

```

### Statistic

5.  Write a null hypothesis and an alternative hypothesis for each question we are asking. (2 points)

Is there a significant difference in the [power output]{.underline} from turbines from the different companies?

-   **Null Hypothesis** (H~0~):\
-   **Alternative Hypothesis** (H~A~):

Is there a significant difference in the required [wind speeds]{.underline} for turbines from different companies?

-   **Null Hypothesis** (H~0~):\
-   **Alternative Hypothesis** (H~A~):

6.  Based on the mean values in the `turbine_summary` data frame and the plots you've created above, [predict the outcome]{.underline} of each t-test that we will run: one comparing wind speeds between turbine companies and another comparing power outputs between companies (graded for completion, not accuracy; 2 points).

    Explain your reasoning (1-2 sentences for each t-test is fine).

    *Answer:*

7.  Perform a t-test on the [power output]{.underline} by turbine maker. (1 point)

```{r}

```

8.  In 2-3 sentences, interpret the output from question 7. Focus on what the p-value is in reference to the cutoff of 0.05, what that means, and whether that means we reject or fail to reject the null hypothesis. (2 points)

    *Answer:*

9.  Perform another t-test, this time on the [wind speed]{.underline} variable by manufacturer. (1 point)

```{r}

```

10. In 2-3 sentences, interpret the output from question 9 (focus on the same ideas as question 8). (2 points)

    *Answer:*
