---
title: '4: Exploring `geom` functions'
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plotting in `ggplot2`

## Data Visualization Types and When to Use Them

### Scavenger Hunt

Head over to [data-to-viz.com](https://www.data-to-viz.com/) and do some exploration! I'll give your group a certain type of plot, and I want you to hunt down as many locations in the flow chart where your plot type is found as you can find.

Once you've done so, click on your plot type and read more about it. Check out the examples that were made in R, including many using `ggplot2`. Take note of the following:

-   which `geom` function is being used to create the plot
-   you'll see a number of other things in the `ggplot` code, some which will be new. What are they? Any idea what they are doing?

## Using `geom` Functions to Make Plots

Let's practice making different kinds of plots with various `geom` functions to see how they work.

```{r}
library(tidyverse)
collars <- read_csv("../data/collar_data.csv")
```

### Histograms

As we learned last modules, histograms are plots which let us look at *one continuous variable* and get a feel for the distribution of that data. To make histograms in `ggplot2`, we use the `geom_histogram()` function. Let's look at signal distance.

```{r}
ggplot(collars, aes(signal_distance)) +
  geom_histogram(bins = 10)

# we can change the number of bins (essentially, the number of columns) by modifying the bins argument in the geom_histogram function
```

### Density plots

Density plots are similar to histograms in that they show the distribution of a *continuous variable* but they are smoothed and continuous version estimated from the data. For this class, you can think of them as being basically the same thing (and use them interchangeably).

```{r}
ggplot(collars, aes(signal_distance)) +
  geom_density()
```

What if we wanted to create a histogram for signal distance for each collar manufacturer? In base R, we would need to filter or make a new dataframe. With `ggplot2`, we can specify that we want the `color` (or the `fill`) to be determined by the maker column.

```{r}
ggplot(collars, aes(signal_distance, color = maker, fill = maker)) +
  geom_density()
```

Hooray! But...we can't really see the bars for Budget Collars LLC in the lower range of the values because they are hidden behind the Collarium Inc. bars. How might we address this?

To fix this issue, we can use the `alpha` argument, which changes the transparency of the bars. The scale for `alpha` runs from 0 (completely transparent) to 1 (completely solid). Let's try 0.5.

```{r}
ggplot(collars, aes(signal_distance, fill = maker)) +
  geom_density(alpha = .5)
```

------------------------------------------------------------------------

#### The powerful and pesky `aes()` function

A quick note about the `aes()` function. It's one of the more confusing bits of `ggplot2`.

When do I put the `color` (or `size` or `linetype` or `fill` or whatever) inside the `aes()` function versus in the `geom` function but outside of `aes()`? When we made our density plots, why did the `color` argument go inside of `aes()` but `alpha` went outside?

Essentially, it boils down to this:

-   if you want something (color, size, etc.) on your plot to change based on a **variable** from a data frame, you will want to put the argument *within* the `aes()` function.

-   if you want something (color, size, etc.) on the plot to be **constant**, you will specify it *outside* of the function.

For some additional examples and explanation, check out [this Stack Overflow page](https://stackoverflow.com/questions/41863049/when-does-the-argument-go-inside-or-outside-aes).

------------------------------------------------------------------------

### Scatter Plot

As a reminder, we use the `geom_point()` function to make a scatter plot of the relationship between *two continuous variables*.

```{r}
ggplot(collars, aes(battery_life, signal_distance)) +
  geom_point()
```

### Bar plot

Let's make a bar plot! They are typically used to represent the counts of *one categorical* variable or the relationship between *one categorical and one continuous variable*. We use the `geom_col` function.

For our example, we can make a bar plot of the number of collars which failed and which did not.

```{r}

fails <- collars %>% 
  group_by(maker, fail) %>% 
  count()
fails
  
ggplot(fails, aes(fail, n, fill = maker)) +
  geom_col()
```

What we just made is something called a "stacked" bar plot, meaning the values for the two collar makers are stacked one on top of the other. If we wanted them side-by-side, we add the argument `position = "dodge"`.

```{r}
ggplot(fails, aes(fail, n, fill = maker)) +
  geom_col(position = "dodge")
```

## Practice

Hopefully you're excited to keep using `ggplot2` to make clean, interpretable data visualizations! With the remaining time in class, spend some time trying to recreate the plots you made in the Module2_1\_DescriptiveStats lesson. Try to recreate both the histogram and the scatter plot.

```{r}
# load the tidyverse package and the penguins package

```

## Summary

Let's summarize some of what we've learned in this module

### Data Visualization Types and When to Use Them

Histogram or Density Plot:

-   one continuous variable

Multiple Histogram or Multiple Density Plot:

-   one continuous variable and one categorical variable

Scatter Plot:

-   two continuous variables
-   can add in a categorical variable as well, but the main relationship is between the two continuous variables

Bar Plot:

-   one continuous variable
-   one continuous and one categorical

Box Plot:

-   one continuous and at least one categorical
