---
title: "Module2_3_SickFish"
author: "Keaton Wilson, EKB"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario

Here is our current scenario.

**What we know**: We found a correlation between eating fish and people
getting sick, but that isn't really the root of the problem. We have
also found that a number of tanks are below the average temperature that
might cause issues with fish immune systems.

**What we don't know**: We don't actually know if rates of disease are
present above average levels in the tanks, and we definitely don't know
what kinds of factors are contributing to this problem (if there is
one!).

Our first challenge is that we can't sample the disease rate in all
tanks. It's too expensive and takes too long. Instead of sampling all
1000 tanks, we've tasked our aquaculture scientists to take a sub-sample
(50 tanks). Our aquaculture scientists have shared this data with us.

Let's scope it out:

```{r}
library(tidyverse)

sick_fish <- read_csv("../data/fish_sick_data.csv")
head(sick_fish)
```

Oh no, we might need to add a Fahrenheit column...let's practice.

```{r}
c_to_f <- function(c = NULL){
  f <- (c * (9/5)) + 32
  return(f)
}

sick_fish <- sick_fish %>% 
  mutate(avg_daily_tempF = c_to_f(avg_daily_temp))
```

## Data Exploration

One of the best ways to get an idea of what our data look like and are
telling us is by calculating some summary statistics.

### Group Challenge

Let's use the `tidyverse` to calculate some summary statistics for the
data we have been given.

Write some code that produces the following values *per species*.

-   mean (average) number of sick fish
-   mean (average) percentage of sick fish per tank
-   how many tanks

```{r}
sick_fish %>%
  group_by(species) %>%
  summarize(mean_sick = mean(num_sick),
            mean_perc = mean(num_sick/num_fish), 
            n = n()) # counts
```

Summary statistics (such as means) are helpful for giving us some ideas
about our data, but they don't tell us the full story. Plotting data can
give us some additional insights.

Before we plot, let's create two different data frames: one for tilapia
and one for trout.

```{r}
sick_tilapia <- sick_fish %>% 
  filter(species == "tilapia")
sick_trout <- sick_fish %>% 
  filter(species == "trout")
```

Now, let's plot our data to get an idea of the distribution (spread) of
the data. We can make histograms to see if there are some tanks that
have a lot of sick fish and are increasing this average or if most of
the tanks seem to have about the same number of sick fish.

What would these two different scenarios tell us?

```{r}
hist(sick_tilapia$num_sick)
hist(sick_trout$num_sick)
```

## Is it temperature?

Does it seem like temperature is the issue? Let's compare values from
cold tanks to our overall values to see if there are major difference or
we can figure out some answers.

```{r}
sick_cold_tilapia <- sick_tilapia %>% 
  filter(avg_daily_tempF < 75)
sick_cold_trout <- sick_trout %>% 
  filter(avg_daily_tempF < 59)

mean(sick_tilapia$num_sick)
mean(sick_cold_tilapia$num_sick)

mean(sick_trout$num_sick)
mean(sick_cold_trout$num_sick)

hist(sick_cold_tilapia$num_sick)
hist(sick_cold_trout$num_sick)
```
