---
title: "Module 2: Food Gone Bad"
author: "Keaton Wilson, modified by EKB"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to Plotting

## Student Learning Outcomes

1.  Students will be able to apply basic data science knowledge to find
    the cause of a real-world scenario--food poisoning!
2.  Students will be able to generate two types of plots using base R
    syntax to visualize a single continuous variable (histograms and
    distributions) and two continous variables (scatterplots).
3.  Students will be able to use visual-thinking skills to create
    visualizations that allow them to explore patterns in data, draw
    inferences, and create solutions.

## Introduction to the problem

We have a wave of people getting sick across the team. People are coming
in complaining of stomach sickness. Doctors have ruled out a
communicable viral infection like norovirus, so it seems likely to be a
food contamination issue.

The two main sources of food that are grown on site and distributed to
team members are [plants grown in hydroponic
greenhouses](https://thespoon.tech/scientists-in-antarctica-grow-food-without-soil-or-sunlight-get-ready-mars/)
(mostly swiss chard, cucumbers and radishes) and fish (tilapia, a
tolerant warm-water species, and rainbow trout, a cold-water species).
Team members vary in the composition of their diet; people are allowed
to choose how much of different food sources they eat.

Fortunately, we have some data to work with! We have data on the
following:

-   which team members are sick
-   how much fish or plant material they incorporate into their diets

### Group Discussion

How might we go about trying to figure out what is causing the problem?

Spend 5 minutes brainstorming in your groups how you might figure out
whether plants or fish are the culprits? Be ready to report out.

## The Data

First we're going to pull in the data and give it a quick
inspection/exploration before we start to work through some of the
visualization tools in R.

```{r}
library(tidyverse)

sick <- read_csv("../data/sick_data.csv")
head(sick)
```

## Visualization in (base) R

Later in the semester, we will use a package from the `tidyverse` called
`ggplot2` which is a very powerful way to make effective visualizations
of data. However, it takes a bit of learning to make even a simple plot.

For now, we are going to use some base R plotting functions to quickly
make some plots.

Data visualization is one of many techniques used to help us understand
our data. One such plot is a histogram.

```{r}
hist(sick$age)
```

### Group Discussion

Discuss this histogram with your group members. Some questions to
consider:

-   What *is* a histogram?
-   What does a histogram tell us?
-   What does each axis mean? (x-axis is horizontal, y-axis is vertical)
-   What conclusions can we draw from this histogram in regards to age
    and sickness?
-   How can we improve this visualization?

Take about 10 minutes. Be ready to report out.

### Group Project

Let's make a better histogram.

Work in your groups to build a histogram of weights with properly
labeled axes, a vertical line that depicts the median, and an
appropriate visually appealing bin size. Check out the `hist()` help
page, and the `abline()` function.

```{r}
hist(sick$age, breaks = 15, xlab = "Age", ylab = "Frequency", main = NULL)
abline(v = median(sick$age))

```

### Group Brainstorm

We've covered one type of visualization that just shows one variable...
but what we're really interested in is figuring out if fish or plants
are the culprit in food poisoning. Spend 10 minutes in your groups
sketching out a visualization that might give us insight into this.

## Scatter plots

```{r}
plot(x = sick$age, y = sick$perc_fish)

# we can also use xlab and ylab here
```

Not super informative, because there isn't really a relationship between
a person's age and the amount of fish they consume. At least not in this
sample.

### Building and Interpreting

In your group, create a scatter plot to determine whether there is a
correlation between the percentage of fish eaten and the number of trips
to the doctor in the past 6 months.

Add some clearer x-axis labels and y-axis labels and aesthetic to make
it as visually appealing as possible. [This
website](http://www.sthda.com/english/wiki/r-plot-pch-symbols-the-different-point-shapes-available-in-r)
will give you some insights into arguments (and values) you can use to
make your plot "prettier." Feel free to have some fun experimenting!

```{r}
plot(x = sick$perc_fish, y = sick$doctor_trips, 
     xlab = "Percentage Fish in Diet", 
     ylab = "Number of Trips to Doctor", 
     pch = 16)

```
