---
title: "Module2_3_SickFish"
author: "Keaton Wilson, EKB"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario

Here is our current scenario.

**What we know**: We found a correlation between eating fish and people
getting sick, but that isn't really the root of the problem. We have
also found that a number of tanks are below the average temperature that
might cause issues with fish immune systems.

**What we don't know**: We don't actually know if rates of disease are
present above average levels in the tanks, and we definitely don't know
what kinds of factors are contributing to this problem (if there is
one!).

Our first challenge is that we can't sample the disease rate in all
tanks. It's too expensive and takes too long. Instead of sampling all
1000 tanks, we've tasked our aquaculture scientists to take a sub-sample
(50 tanks). Our aquaculture scientists have shared this data with us.

Let's scope it out:

```{r}

```

Oh no, we might need to add a Fahrenheit column...let's practice.

```{r}
c_to_f <- function(c = NULL){
  f <- (c * (9/5)) + 32
  return(f)
}

```

## Data Exploration

One of the best ways to get an idea of what our data look like and are
telling us is by calculating some summary statistics.

### Group Challenge

Let's use the `tidyverse` to calculate some summary statistics for the
data we have been given.

Write some code that produces the following values *per species*.

-   mean (average) number of sick fish
-   mean (average) percentage of sick fish per tank
-   how many tanks

```{r}

```

Summary statistics (such as means) are helpful for giving us some ideas
about our data, but they don't tell us the full story. Plotting data can
give us some additional insights.

Before we plot, let's create two different data frames: one for tilapia
and one for trout.

```{r}

```

Now, let's plot our data to get an idea of the distribution (spread) of
the data. What might the distribution of the data tell us that an
average can't?

We can make histograms to see if there are some tanks that have a lot of
sick fish and are increasing this average or if most of the tanks seem
to have about the same number of sick fish.

What would these two different scenarios tell us?

```{r}

```

Let's compare values from tanks which are below the temperature cutoffs
to those which are above the cutoffs to see if there are major
difference or we can figure out some answers.

## Using the `if_else()` function

Before we start calculating means and plotting, let's create a new
column in our data frame to indicate whether the tank temperature is
above or below the cutoff temperature. We will use our fish-specific
data frames to do this.

A useful function that we can use when we want to create a new column
based on values in another column is the `if_else()` function. It
operates the following way: 1. IF a condition is true, do [something].
2. IF a condition is false, do [something ELSE].

```{r}

```

### Individual Challenge

Practice using the `if_else()` function as we did above. This time, use
the sick_trout dataframe. Remember to change the bits of the code that
you need to!

```{r}

```

## Another Type of Iteration: `for loop`

Another way we could write something like this is by using the
`if_else()` function in something we call a `for loop`.

Before we do that, though, let's talk through the general structure of a
`for loop`. It essentially says for each value in a list, do a certain
task. The "loop" is because we are "looping" through a list of values,
performing the task for one value then looping back to the beginning to
perform the task for the next value. We type the "task" within curly
brackets, similar to a function that we write.

```{r}

```

Now that we know the general structure of a `for loop`, we can combine
it with the `if_else()` function to create a new column.

```{r}

```

*Note: I will never ask you to write a for loop completely from scratch.
I might have you copy and paste one or change some values in one, but
you won't have to write one out yourself.*

### Group Challenge

Try your hand at using the `for loop` we wrote above to create a new
temperature cutoff column in the `sick_trout` data frame. Remember, the
cutoff for trout was 59 F.

```{r}

```

### Why `for loops`?

Like with our last lesson about functions, I've asked you to perform a
task in a new and complicated way than you need to for that task. Why?

You'll find some examples
[here](https://github.com/bleds22e/PP_shifts/blob/master/scripts/functions.R)
in code written for my PhD dissertation.

## Back to Data Exploration

We now have 2 data frames, one with tilapia data and one with trout
data. Each data frame also has a new column called `temp_cutoff`. On
your own or with a partner, start exploring the data to figure out if
there are differences between warm and cold tilapia and warm and cold
trout.

Can you pinpoint an issue? Let's start by comparing means.

```{r}

```

Not much popping out in the means. Next thing to check would be
histograms of the number of sick fish for both species, above and below
the cutoffs.

For now, I would recommend making 4 different data frames (this isn't
"best practice" but it is really helpful while you are learning).

```{r}

```
