---
title: "Module 2 Wrap-Up"
author: "Ellen Bledsoe"
date: "2023-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is causing the food poisoning?

At the beginning of Module 2, we set out to discover what was causing food poisoning among our colleagues at our Antarctic base. Let's put everything we've learned about descriptive statistics and data visualization to use to try to hunt down what the problem is.

## Set-up

Let's load the package and data we will need.

### Packages

```{r message = FALSE}

```

### Data

First, we need our dataset!

```{r}

```

Let's check out our data and remind ourselves what we are working with.

```{r}

```

## Which Fish?

Last class, we plotted the number of sick fish. Let's remind ourselves what that looked like. Make a plot that compares the numbers of sick fish per species. We actually have a few options!

```{r}

```

### Density

Wait a second! Take a look back at the data. There is a "number of fish" column, indicating the total number of fish in the tank, and it looks like those numbers can differ pretty widely.

We should probably take into account how many fish there are in the tank to begin with. 12 sick fish out of 50 is probably a bigger deal than 12 sick fish out of 100!

What we need to do is calculate a density of fish--number of sick fish / number of total fish.

```{r}
# take into account the number of fish in the tank: density of sick fish

```

#### Small Groups

Let's make sure our conclusions about trout being the true culprits still hold when we account for the total number of fish in the tank.

Work in small groups to do the following:

-   find the average number and standard deviation of sick fish for both species
-   make a plot that compares the distributions of sick fish numbers for both species (you have multiple options here!)

```{r}


```

Uh oh...the trout densities look even worse than just the number of sick fish. We need to take a closer look at what is going on in the trout tanks!

We should create a data frame that only contains trout to work with for the rest of our analyses. Take a few minutes to work on that; call it `sick_trout`.

```{r}

```

## What Environmental Factor?

Take a look back at the data frame. Which columns are environmental variables that could be driving the issues?

Are those columns continuous or categorical? What plot type have we talked about that might help us find a relationship between density and each of these variables (one at a time...)?

In small groups, make plots using `density` column in the `sick_trout` data to try to figure out which environmental factor is causing problems in the trout. Treat your variables as continuous.

```{r}

```

What do we think is the environmental driver causing issues with the trout?
